### 2018-06-29 pubs/itp-2018/sections/proof-reconstruction/canonicalize.tex
Fixies of the day

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:canon-or}
  \begin{aligned}
   &\hspace{.495mm}\fcanon_{∨} : \NProp \to \NProp\\
   &\begin{array}{lll}
     \fcanon_{∨} &(⊥ ∨ φ)     &= \fcanon_{∨}~φ \\
     \fcanon_{∨} &(φ ∨ ⊥)     &= \fcanon_{∨}~φ \\
     \fcanon_{∨} &(⊤ ∨ φ)     &= ⊤\\
     \fcanon_{∨} &(φ ∨ ⊤)     &= ⊤\\
     \fcanon_{∨} &(φ₁ ∨ φ₂)   &=
      \begin{cases}
       ⊤,                     & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∨} φ₂;\\
       ⊤,                     & \text{ if } (¬ φ₁) ∈_{∨} φ₂;\\
       \fcanon_{∨}~φ₂,        & \text{ if } φ₁ ∈_{∨} φ₂;\\
       ⊤,                     & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊤;\\
       φ₁,                    & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊥;\\
       φ₁ ∨ \fcanon_{∨}~φ₂,   & \text{ otherwise.}
      \end{cases}\\
     \fcanon_{∨} &φ           &= φ.
    \end{array}
  \end{aligned}
\end{empheq}

@@

  \begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
   \label{eq:simplify-and}
    \begin{aligned}
     &\hspace{.495mm}\fcanon_{∧} : \NProp \to \NProp\\
      &\begin{array}{lll}
        \fcanon_{∧} &(⊥ ∧ φ)     &= ⊥  \\
        \fcanon_{∧} &(φ ∧ ⊥)     &= ⊥  \\
        \fcanon_{∧} &(⊤ ∧ φ)     &= \fcanon_{∧}~φ \\
        \fcanon_{∧} &(φ ∧ ⊤)     &= \fcanon_{∧}~φ \\
        \fcanon_{∧} &(φ₁ ∧ φ₂) &=
          \begin{cases}
            ⊥,                   & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∧} φ₂;\\
            ⊥,                   & \text{ if } (¬ φ₁) ∈_{∧} φ₂;\\
            \fcanon_{∧}~φ₂,      & \text{ if } φ₁ ∈_{∧} φ₂;\\
            φ₁,                  & \text{ if } \fcanon_{∧}~φ₂~≡ ⊤;\\
            ⊥,                   & \text{ if } \fcanon_{∧}~φ₂~≡ ⊥;\\
            φ₁ ∧ \fcanon_{∧}~φ₂, &\text{ otherwise.}
          \end{cases}\\
        \fcanon_{∧} &φ         &= φ.
       \end{array}
    \end{aligned}
    \end{empheq}
\end{mainlemma}

Using the functions $\fcanon_{∨}$ and $\fcanon_{∧}$ we can define our function for
the negative normal form present in \Metis, we once again formulate this version
based on the \Metis source code to \emph{normalize} formulas. We use bounded
recursion to define the function $\fnnf_{1}$ in~\eqref{eq:nnf}.

@@

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:nnf}
  \begin{aligned}
    &\hspace{.495mm}\fnnf_{1} : \Prop \to \Bound \to \Prop\\
    &\begin{array}{llll}
    \fnnf_{1} &(φ₁ ∧ φ₂)     &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~φ₁~n ∧ \fnnf_{1}~φ₂~n))\\
    \fnnf_{1} &(φ₁ ∨ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~φ₁~n ∨ \fnnf_{1}~φ₂~n))\\
    \fnnf_{1} &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ φ₂)~n))\\
    \fnnf_{1} &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ (¬ φ₂))~n))\\
    \fnnf_{1} &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₁) ∧ (¬ φ₂))~n))\\
    \fnnf_{1} &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₂) ∧ φ₁)~n))\\
    \fnnf_{1} &(¬ (¬ φ))     &(\suc~n) &= \fnnf_{1}~φ₁~n\\
    \fnnf_{1} &(¬ ⊤)         &(\suc~n) &= ⊥\\
    \fnnf_{1} &(¬ ⊥)         &(\suc~n) &= ⊤\\
    \fnnf_{1} &φ             &\zero    &= φ
    \end{array}
  \end{aligned}
\end{empheq}

The function $\fnnf$ in~\eqref{eq:fnnf} is defined similarly as the \fstrip
function in~\eqref{eq:strip} with a function $\fnnf_{cm}$ that finds a bound
for its recursive calls (See its definition in Appendix in~\cite{Prieto-Cubides2017a}).

>>>>

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:canon-or}
  \begin{aligned}
   &\hspace{.495mm}\fcanon_{∨} : \NProp \to \NProp\\
   &\begin{array}{lll}
     \fcanon_{∨} &(⊥ ∨ φ)     &= \fcanon_{∨}~φ \\
     \fcanon_{∨} &(φ ∨ ⊥)     &= \fcanon_{∨}~φ \\
     \fcanon_{∨} &(⊤ ∨ φ)     &= ⊤\\
     \fcanon_{∨} &(φ ∨ ⊤)     &= ⊤\\
     \fcanon_{∨} &(φ₁ ∨ φ₂)   &=
      \begin{cases}
       ⊤,                     & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∨} φ₂;\\
       ⊤,                     & \text{ if } (¬ φ₁) ∈_{∨} φ₂;\\
       \fcanon_{∨}~φ₂,        & \text{ if } φ₁ ∈_{∨} φ₂;\\
       ⊤,                     & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊤;\\
       φ₁,                    & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊥;\\
       φ₁ ∨ \fcanon_{∨}~φ₂,   & \text{ otherwise,}
      \end{cases}\\
     \fcanon_{∨} &φ           &= φ.
    \end{array}
  \end{aligned}
\end{empheq}

@@

  \begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
   \label{eq:simplify-and}
    \begin{aligned}
     &\hspace{.495mm}\fcanon_{∧} : \NProp \to \NProp\\
      &\begin{array}{lll}
        \fcanon_{∧} &(⊥ ∧ φ)     &= ⊥  \\
        \fcanon_{∧} &(φ ∧ ⊥)     &= ⊥  \\
        \fcanon_{∧} &(⊤ ∧ φ)     &= \fcanon_{∧}~φ \\
        \fcanon_{∧} &(φ ∧ ⊤)     &= \fcanon_{∧}~φ \\
        \fcanon_{∧} &(φ₁ ∧ φ₂) &=
          \begin{cases}
            ⊥,                   & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∧} φ₂;\\
            ⊥,                   & \text{ if } (¬ φ₁) ∈_{∧} φ₂;\\
            \fcanon_{∧}~φ₂,      & \text{ if } φ₁ ∈_{∧} φ₂;\\
            φ₁,                  & \text{ if } \fcanon_{∧}~φ₂~≡ ⊤;\\
            ⊥,                   & \text{ if } \fcanon_{∧}~φ₂~≡ ⊥;\\
            φ₁ ∧ \fcanon_{∧}~φ₂, &\text{ otherwise,}
          \end{cases}\\
        \fcanon_{∧} &φ         &= φ.
       \end{array}
    \end{aligned}
    \end{empheq}
\end{mainlemma}

Using the functions $\fcanon_{∨}$ and $\fcanon_{∧}$ we formulate the functions
that finds the normal form based on the \Metis source code used to
\emph{normalize} formulas. We use bounded recursion to define the function
$\fnnf_{1}$ in~\eqref{eq:nnf}.

@@

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:nnf}
  \begin{aligned}
    &\hspace{.495mm}\fnnf_{1} : \Prop \to \Bound \to \Prop\\
    &\begin{array}{llll}
    \fnnf_{1} &(φ₁ ∧ φ₂)     &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~φ₁~n ∧ \fnnf_{1}~φ₂~n))\\
    \fnnf_{1} &(φ₁ ∨ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~φ₁~n ∨ \fnnf_{1}~φ₂~n))\\
    \fnnf_{1} &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ φ₂)~n))\\
    \fnnf_{1} &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ (¬ φ₂))~n))\\
    \fnnf_{1} &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₁) ∧ (¬ φ₂))~n))\\
    \fnnf_{1} &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₂) ∧ φ₁)~n))\\
    \fnnf_{1} &(¬ (¬ φ))     &(\suc~n) &= \fnnf_{1}~φ₁~n\\
    \fnnf_{1} &(¬ ⊤)         &(\suc~n) &= ⊥\\
    \fnnf_{1} &(¬ ⊥)         &(\suc~n) &= ⊤\\
    \fnnf_{1} &φ             &\zero    &= φ.
    \end{array}
  \end{aligned}
\end{empheq}

The function $\fnnf$ in~\eqref{eq:fnnf} is defined by using the function $\fnnf_{cm}$ encharges to find
the bound for the recursive calls of the $\fnnf_{1}$ function. The interested reader can find the complete
definition from Appendix in~\cite{Prieto-Cubides2017a}, we do not include it for the seek of brevity.

### 2018-06-29 pubs/itp-2018/sections/proof-reconstruction/resolve.tex
Nice box color for equations about functions
Fixies of the day

The \resolve inference rule in combination with the \simplify rule are the
responsable rules to perform resolution. The \fresolve rules is a particular
version of the resolution theorem for propositional logic \TSTP derivations (see
this rule in~Fig.~\ref{fig:metis-inferences}).

\begin{example} In the \TSTP derivation in
Fig.~\ref{eq:complete-example-problem}, there are two ocurrences in
line~\ref{line14} and line~\ref{line16}~of the \resolve rule. This rule has two
inputs: the \emph{resolvent}, a literal denoted by ℓ in~Fig.\ref{fig:metis-inferences},
and a list of two formulas. \end{example}

In~\cite{Prieto-Cubides2017a}, we describe our first attempt to reconstruct this
rule that performs reordering of formulas jointly with the application of a
customized version of the resolution theorem to reconstruct the rule. We propose
an alternative definition formulated as a particular case
of the \fsimplify rule. We do this after seeing the behavior
of this rule in the \TSTP derivations from a more practical point of view for
the type-checking.

\begin{mainth}
  \label{thm:resolve}
  Let $ℓ$ be a literal, $ℓ : \Lit$, and $ψ : \Target$. If $Γ ⊢ φ₁$ and
  $Γ~⊢~φ₂$ then $Γ~⊢~\fresolve~φ₁~φ₂~ℓ~ψ$, where
  \begin{equation}
  \begin{split}
  \label{eq:resolve}
    &\fresolve : \Source \to \Source \to \Lit \to \Target \to \Prop\\
    &\fresolve~φ₁~φ₂~ℓ~ψ~=~\fsimplify~(φ₁ ∧ φ₂)~(¬ ℓ ∨ ℓ)~ψ
  \end{split}
  \end{equation}
\end{mainth}

>>>>

The \resolve inference rule in combination with the \simplify rule are the
responsable rules to perform resolution, in particular, the \resolve rule is is
the version of the resolution theorem showed in~Fig.~\ref{fig:metis-inferences}.

\begin{example} In the \TSTP derivation in
Fig.~\ref{eq:complete-example-problem}, there are two ocurrences in
line~\ref{line14} and line~\ref{line16}~of the \resolve rule. This rule has two
inputs in the following order: the \emph{resolvent} which is the literal denoted
by ℓ in~Fig.~\ref{fig:metis-inferences} and a list formed by two formulas to
resolve. \end{example}

In~\cite{Prieto-Cubides2017a}, we describe our first attempt to reconstruct this
rule that performs reordering of formulas jointly with the application of a
customized version of the resolution theorem to reconstruct the rule.
Unfortunately, that approach suffers of an exponential computational cost for
each searching by reordering the formulas. This time, we propose an alternative
formulation to reconstruct this rule as a particular case of the \fsimplify rule.
This observation was taken from the \TSTP derivations in order to improve
the timing of the type-checking. Futher improvements are future work.

\begin{mainth}
  \label{thm:resolve}
  Let $ℓ$ be a literal, $ℓ : \Lit$, and $ψ : \Target$. If $Γ ⊢ φ₁$ and
  $Γ~⊢~φ₂$ then $Γ~⊢~\fresolve~φ₁~φ₂~ℓ~ψ$, where
  \begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \begin{split}
  \label{eq:resolve}
    &\fresolve : \Source \to \Source \to \Lit \to \Target \to \Prop\\
    &\fresolve~φ₁~φ₂~ℓ~ψ~=~\fsimplify~(φ₁ ∧ φ₂)~(¬ ℓ ∨ ℓ)~ψ.
  \end{split}
  \end{empheq}
\end{mainth}
